---
title: "Creating Interactive Line Plots Using Plotly"
author: "Giang Rudderham, University of Iowa"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Development Notes

These are notes for yourself while you're working on your tutorial - remove them later.

## Topic

What's your topic?

* Language or Program: R
* Tutorial Type (Task, Technique, Package Overview, Tool): Task
* Specific Package/Task/Technique/Tool: Creating interactive line plots using plotly


# Introduction

In this tutorial, we will cover how to create an interactive line plot using the package `plotly` in R.

## Setup

The package `plotly` allows us to create interactive data visualizations with no required knowledge of web publishing. 

For this tutorial, we will use the polling data for the presidential election 2020 from fivethirtyeight.com

First, load the `tidyverse` and `plotly` packages. We will use these two packages to clean and shape the data and then create the visualization.

Then read the dataset as a csv.

```{r}
library(plotly)
library(tidyverse)

df <- read.csv("president_polls.csv", stringsAsFactors = F)

```


# Section 1: Filter to keep only the data we need  

First, let's keep only the candidates from the 2 major political parties.

We will use the `filter` function in the package `tidyverse` to filter and keep those 2 candidates.

Make sure you have run the code in the "Set Up" section above.

```{r}
# filter to keep only Joe Biden and Donald Trump
df_final_candidates <- filter(df, candidate_id %in% c(13256, 13254))

```


## EXERCISE 1

Five Thirty Eight rates the quality of each pollster on a scale of A (best) to D- (worst).

Using the `filter` function, get rid of pollsters that have D- rating by Five Thirty Eight.

Make sure you have run the code in the "Set Up" section above.

```{r}
# filter to get rid of pollsters that have D- rating by Five Thirty Eight
df_final_candidates <- filter(df_final_candidates, fte_grade != "D-")
```


# Section 2: Convert date variables to date format

In order to use a date variable to create a data visualization, we first need to make sure that the variable has date format.

For example, here we convert the `start_date` variable to a date format:

```{r}
# convert start_date to date format
df_final_candidates <- mutate(df_final_candidates, start_date_fmt = as.Date(start_date, format = "%m/%d/%y"))
              
```

## EXERCISE 2

Using the `as.Date` function, convert the variable `end_date` to a date format.

```{r}
# convert end_date to date format
df_final_candidates <- mutate(df_final_candidates, end_date_fmt = as.Date(end_date, format = "%m/%d/%y"))
```


# Section 3: Get Averages for Each Date

```{r}
# group by and summarise to get averages of all polls for each end date
df_plot <- df_final_candidates %>%
  group_by(end_date_fmt, answer) %>%
  summarise(mean_pct = mean(pct, na.rm = T))
  
```

# Section 4: Calculate Rolling Averages

```{r}
# calculate 14-day simple rolling averages
df_plot <- df_plot %>%
  group_by(answer) %>%
  mutate(rolling_mean_pct = zoo::rollmean(mean_pct, k = 14, fill = NA, align = "right")) %>%
  ungroup()
```
               
# Section 5: Create the Plot

```{r}
# create color palette 
pal <- c("#0571b0", "#ca0020")

# create the plot
p <- plot_ly(df_plot, x = ~end_date_fmt, y = ~mean_pct, type = "scatter", mode = "markers", 
             color = ~answer, colors = pal, opacity = 0.25, hoverinfo = "none", showlegend = F) %>%
  add_trace(y = ~rolling_mean_pct, type = "scatter", mode = "lines", color = ~answer, colors = pal, 
            opacity = 1, hoverinfo = "all", hovertemplate = paste0("%{x}:<br>", "%{y:.1f}%"),
            showlegend = T) %>%
  layout(title = "<b>Presidential General Election Polls</b><br>14-Day SMA. Data Updated Oct 7, 2020",
         xaxis = list(title = ""),
         yaxis = list(title = ""),
         hovermode = "x unified")
  
p
```



# Wrap-up

Remind folks of what they've learned.

Where can they learn more?




